---
title: "HW8_490IDS_12"
output: html_document
date: "`r format(Sys.time(), '%B %d, %Y')`"
---

###Exercises

***

This assignment will use Google Earth for data display. The .rda file is uploaded to Moodle. 
Load HW8.rda and attach the XML library

```{r}
library("XML")
```

##Part 1.  Create the data frame
Look at the instructions in HW8.pdf. 
Functions you'll want to use: xmlParse(), xmlRoot(), xpathSApply(), xmlGetAttr().
It also might make it easier to use: xmlToList(), merge().

Load the data frame called LatLon from HW8.rda.  

```{r}
load("/Users/mikedu/Desktop/490/hw8.rda")
dim(LatLon)
names(LatLon)
```

Download the gzipped XML factbook document from http://jmatchparser.sourceforge.net/factbook/ and create an XML "tree" in R 

```{r}
factbook_doc <- xmlParse("file:///Users/mikedu/Desktop/490/factbook.xml")
factbook_root <- xmlRoot(factbook_doc)
class(factbook_root)
```

Use XPath to extract the infant mortality and the CIA country codes from the XML tree
Create a data frame called IM using this XML file.
The data frame should have 2 columns: for Infant Mortality and CIA.Codes.

```{r}
path = "//field[@name = 'Infant mortality rate']/rank"
Infant.Mortality <- xpathSApply(factbook_root, path, xmlGetAttr, "number")
CIA.Codes <- xpathSApply(factbook_root, path, xmlGetAttr, "country")
IM <- data.frame("Infant.Mortality" = as.numeric(Infant.Mortality), "CIA.Codes" = CIA.Codes)
```

Extract the country populations from the same XML document
Create a data frame called Pop using these data.
This data frame should also have 2 columns, for Population and CIA.Codes.

```{r}
path1 = "//field[@name = 'Population']/rank"
Population <- xpathSApply(factbook_root, path1, xmlGetAttr, "number")
CIA.Codes <- xpathSApply(factbook_root, path1, xmlGetAttr, "country")
Pop <- data.frame("Population" = as.numeric(Population), "CIA.Codes" = CIA.Codes)
```

Merge the two data frames to create a data frame called IMPop with 3 columns: IM, Pop, and CIA.Codes

```{r}
IMPop <- merge(IM, Pop, "CIA.Codes")
IMPop$CIA.Codes <- toupper(IMPop$CIA.Codes)
```

Now merge IMPop with LatLon (from newLatLon.rda) to create a data frame called AllData that has 6 columns for Latitude, Longitude, CIA.Codes, Country Name, Population, and Infant Mortality (please check lat,long are not reversed in the file)

```{r}
AllData <- merge(IMPop, LatLon, "CIA.Codes")
```


***

##Part 2.  Create a KML document
Make the KML document described in HW8.pdf. It should have the basic structure shown in that document. You can use the addPlacemark function below to make the Placemark nodes, you just need to complete the line for the Point node and figure out how to use the function.

```{r}
makeBaseDocument = function(){
### This code creates the template KML document
  doc = newXMLDoc()
  root = newXMLNode("kml", namespaceDefinitions="http://www.opengis.net/kml/2.2", doc = doc)
  Document = newXMLNode("Document", parent = root)
  newXMLNode("Name", "Country Facts", parent = Document)
  newXMLNode("Description", "Infant Mortality", parent = Document)
  LookAt = newXMLNode("LookAt", parent = Document)
  newXMLNode("longitude", "-121", parent = LookAt)
  newXMLNode("latitude", "43", parent = LookAt)
  newXMLNode("altitude", "4100000", parent = LookAt)
  newXMLNode("title", "0", parent = LookAt)
  newXMLNode("heading", "0", parent = LookAt)
  newXMLNode("altitudeMode", "absolute", parent = LookAt)
  Folder = newXMLNode("Folder", parent = Document)
  newXMLNode("Name", "CIA Fact Book", parent = Folder)
  return(doc)
}
```

```{r}
addPlacemark = function(lat, lon, ctryCode, ctryName, pop, infM, parent, 
                        inf1, pop1, style = FALSE)
{
  pm = newXMLNode("Placemark", 
                  newXMLNode("name", ctryName), attrs = c(id = ctryCode), 
                  parent = parent)
  newXMLNode("description", paste(ctryName, "\n Population: ", pop, 
                                  "\n Infant Mortality: ", infM, sep =""),
             parent = pm)
  
  newXMLNode("Point",newXMLNode("coordinates", paste(lat, lon, sep =",")),
             parent = pm)
      
### You need to fill in the code for making the Point node above, including coordinates.
### The line below won't work until you've run the code for the next section to set up
### the styles.

  if(style) newXMLNode("styleUrl", paste("#YOR", inf1, "-", pop1, sep = ''), parent = pm)
}
```

```{r}
doc1 <- makeBaseDocument()
root1 <- xmlRoot(doc1)
DocNode1 <- root1[[1]][4]
for(i in 1:(dim(AllData)[1])){
  addPlacemark(lat = AllData$Latitude[i], lon = AllData$Longitude[i], ctryCode = AllData$CIA.Codes[i], ctryName = AllData$Country.Name[i], 
               pop = AllData$Population[i], infM = AllData$Infant.Mortality[i], parent = DocNode1)
}
```

Save your KML document here, call it Part2.kml, and open it in Google Earth. (You will need to install Google Earth.)  
It should have pushpins for all the countries.

```{r}
saveXML(doc1, "Part2.kml")
```


***

##Part 3.  Add Style to your KML
Now you are going to make the visualizatiion a bit fancier.  Pretty much all the code is given to you below to create style elements that are to be placed near the top of the document. These , you just need to figure out what it all does.

Start fresh with a new KML document, by calling makeBaseDocument()

```{r}
doc2 = makeBaseDocument()
```

The following code is an example of how to create cut points for different categories of infant mortality and population size.
Figure out what cut points you want to use and modify the code to create these categories.

```{r}
infCut = cut(AllData$Infant.Mortality, breaks = c(0, 10, 25, 50, 75, 200))
infCut = as.numeric(infCut)
popCut = cut(log(AllData$Population), breaks = 5)
popCut = as.numeric(popCut)
```

Now figure out how to add styles and placemarks to doc2
You'll want to use the addPlacemark function with style = TRUE

Below is code to make style nodes. 
You should not need to do much to it.

You do want to figure out what scales to you for the sizes of your circles

```{r}
scales = c(0.5, 1, 3, 5, 10)
circleCol=c("blue", "green", "yellow", "orange", "red")

addStyle = function(col1, pop1, parent, urlBase, scales = scales)
{
  st = newXMLNode("Style", attrs = c("id" = paste("YOR", col1, "-", pop1, sep="")), parent = parent)
  newXMLNode("IconStyle", 
			 newXMLNode("scale", scales), 
		     newXMLNode("Icon", paste(urlBase, "color_label_circle_", circleCol[col1], ".png", sep ="")), parent = st)
}
root2 = xmlRoot(doc2)
DocNode2 = root2[1]

for (k in 1:5)
{
  for (j in 1:5)
  {
    addStyle(j, k, DocNode2, 'http://www.stanford.edu/~vcs/StatData/circles/', scales[k])
  }
}
```

You will need to figure out what order to call addStyle() and addPlacemark() so that the tree is built properly. You may need to adjust the code to call the png files

```{r}
DocNode3 = root2[[1]][4]
for(i in 1:(dim(AllData)[1])){
  addPlacemark(lat = AllData$Latitude[i], lon = AllData$Longitude[i], ctryCode = AllData$CIA.Codes[i], ctryName = AllData$Country.Name[i], 
               pop = AllData$Population[i], infM = AllData$Infant.Mortality[i], parent = DocNode3,
               inf1 = infCut[i], pop1 = popCut[i], style = TRUE)
}
```

Finally, save your KML document, call it Part3.kml and open it in Google Earth to  verify that it works.  For this assignment, you only need to submit your code, nothing else.  You can assume that the grader has already loaded HW8.rda.

```{r}
saveXML(doc2, "Part3.kml")
```
